Agda Meets Accelerate

### 2012-08-01 paper/main.tex
Revise related work

A familiar example for an indexed data type is the type
\verb+Vec A n+ for vectors of fixed length \verb+n+ and elements of
type \verb+A+ program program with an access operation that restricts the index to the
actual length of the vector at compile time.\footnote{An
  identifier can be an almost arbitrary 
  string of Unicode characters except spaces, parentheses, and curly
  braces. Agda also supports mixfix syntax with the position of
  arguments indicated by underscores in the defining occurrence of an
  identifier.} 
\begin{verbatim}
data Vec (A : Set) : Nat -> Set where
  []   : Vec A zero
  _::_ : {n : Nat} -> A -> Vec A n -> Vec A (suc n)
\end{verbatim}
The above defines an indexed data type \verb+Vec A n+ with two
constructors, \verb+[]+ for the vector of length zero and 
\verb+_::_+ for the infix cons operator that increases the length by one.

@@

Dependent types for array programming: 
\begin{itemize}
\item "Dependent Types for Distributed Arrays".
Wouter Swierstra and Thorsten Altenkirch. Postproceedings of TFP 2008.
\item "More Dependent Types for Distributed Arrays".
Wouter Swierstra. Accepted for publication in the Journal of Higher-order and Symbolic Computation.
\end{itemize}
This work focuses on describing distribution strategies using dependent types (which is orthogonal to our work). However, to characterise distributions, it also tracks bounds in types.

DML:
\begin{itemize}
\item Xi, Hongwei (March 2007). "Dependent ML: An Approach to Practical Programming with Dependent Types". Journal of Functional Programming 17 (2). [There is a previous ICFP paper co-authored with his advisor Bob Harper.]
\end{itemize}
Extension of ML to track array bounds.

Other embedded array languages:
\begin{itemize}
\item Accelerator: C++ library providing data-parallel GPU programming; subsequent work by Satnam Singh provides an F\# binding (but that was never published as for as I know---double check); the language provides only a small subset of the functionality of Accelerate
\item Copperhead: embedded array language in Python; no static guarantees; performance varied widely in our experiments
\item anything else?
\end{itemize}

@@

%
% ---- Bibliography ----
%
\bibliography{abbrevs,papers,books,collections,misc,theses}
\bibliographystyle{abbrv}

>>>>

A familiar example for an indexed data type is the type
\verb+Vec A n+ for vectors of fixed length \verb+n+ and elements of
type \verb+A+ program program with an access operation that restricts the index to the
actual length of the vector at compile time.\footnote{An
  identifier can be an almost arbitrary 
  string of Unicode characters except spaces, parentheses, and curly
  braces. Agda also supports mixfix syntax with the position of
  arguments indicated by underscores in the defining occurrence of an
  identifier.} 
\begin{verbatim}
data Nat : Set where
  zero : Nat
  suc  : Nat -> Nat

data Vec (A : Set) : Nat -> Set where
  []   : Vec A zero
  _::_ : {n : Nat} -> A -> Vec A n -> Vec A (suc n)
\end{verbatim}
The above defines an indexed data type \verb+Vec A n+ with two
constructors, \verb+[]+ for the vector of length zero and 
\verb+_::_+ for the infix cons operator that increases the length by one.

@@

Peebles formalised parts of the API of the Haskell array library Repa using Agda~\cite{peebles:derpa}. The formalisation uses the same shape structure as we are using in Accelerate, but array computations are neither embedded nor can parallel high-performance code be generated.

Swierstra and Altenkirch investigated the use of dependent types for \emph{distributed} array programming~\cite{swierstra-altenkirch:dep-types-for-distr-arrays,swierstra:more-dependent-types}. Their notation for distributed arrays was inspired by X10 and the main focus is on expressing locality aware algorithms.

Dependent ML is an ML dialect with restricted form of dependent types, which, among other applications, may be used to statically check array bounds in array operations~\cite{xi:dml-jfp}. However, only simple indexing and array updating are considered and not aggregate array operations, such as those provided by Accelerate.

Accelerator~\cite{Tarditi:2006} enables embedded GPU computations in C\# programs; it subsequently also added F\# support. However, no attempt is made to track properties of array programs statically. Similarly, Copperhead~\cite{Catanzaro:EECS-2010-124} embedded an array language into Python, but does not attempt to track information statically.

@@

%
% ---- Bibliography ----
%
\bibliography{local,abbrevs,papers,books,collections,misc,theses}
\bibliographystyle{abbrv}
